---
title: "Flow_Cytometry_Screen Analysis"
author: "Nathan Brandt"
date: "June 19th, 2016"
output: html_document
---
#The final goal of this file is to implement a modular R code for Cytometric Analysis


#1) Two samples in four media types analyzed on the Accuri in triplicate .
#    Strains-
#     Plate 01 of the GAP1::GFP KO Collection Pool
#     Strains have GFP fused with GAP1 and tdTomato expressed under the ACT1 promoter in the HO locus
#    Media Types-
#     SC
#   Controls-
#      FY4 in Proline Limited Media
#      FY4 in SC Media
#      GFP::GAP1 HISKO in Proline Limited Media
#      GFP::GAP1 HISKO in SC Media

###This code is designed for use with the Accuri flow cytometer, which is equiped with the following lasers and filters
* Blue laser (488 nm)
  + FL1 filter = 514/20nm   GFP
  + FL3 filter = 575/25nm   YFP

* Yellow/green laser (552 nm)
  + FL2 filter = 610/20nm   mCherry, dtomato
  + FL4 filter = 586/15nm   DsRed

******
##Step 1: Load relevant libraries 

```{r}
#Load libraries
library(flowCore)
library(flowViz)
library(flowQ)
library(flowStats)
library(Hmisc)
library(reshape2)
library(ggplot2)
library(flowWorkspace)
library(ggcyto)
```

###Step 2: Read in all .fcs files in current directory and a sample sheet that contains four columns with 
* column1 = Well
* column2 = Strain
* column3 = Genotype
* column4 = Ploidy
* column5 = Media


```{r}
#Set working directory to the folder in which you have stored your .fcs files
#Read in all the fcs files in the directory, with alter.ames changing "-" to "."

flowData <- read.flowSet(path = "~/Google Drive/Transfer_Folder/GAP1_GFP_Screen_FSC_Files/Plate01_SC", pattern=".fcs", alter.names=TRUE)

sample.sheet <- read.csv("~/Google Drive/Transfer_Folder/GAP1_GFP_Screen_DataSheets/Plate01_SC.csv")
```

```{r}
#Check how many cells were counted in each fcs file
fsApply(flowData, each_col, length)

#The medians of data
fsApply(flowData, each_col, median)

```


Filter Data
#Filter Data
```{r}
##Apply a rectangle gate based on FSC and SSC to remove values that we have determined are typically debris for haploid cells.  This should be adjusted if the cell type is different
rectGate <- rectangleGate("FSC.A"= c(5.6,7.6),"SSC.A" = c(4.6,6.8))

#Subset the filtered data to remove outliers defined by FFC and SSC 
filterData <- flowData#Subset(flowData, rectGate) 

#Check medians of filtered data
fsApply(filterData, each_col, median)
```

#Plot the rectangle gate used for filtering and determine the proportion of cells used for subsequent analysis
```{r}
##Plot the rectangle gate used to filter on the basis of SSC and FSC to ensure that most cells are retained in analysis.
xyplot(FSC.A ~ SSC.A, data=flowData, ylim=c(4,8),xlim=c(0,8), displayFilter=TRUE, filter=rectGate, smooth=F, xbin=128, stat=T, pos=0.5, abs=T)
```

###Plot an overview of the data
```{r}
##Plot FL1(GFP) versus FSC 
xyplot(FL1.A ~ FSC.A, data=filterData, ylim=c(0,8),xlim=c(4,8),smooth=F, xbin=128)

##Plot FFS versus SSC
xyplot(SSC.A ~ FSC.A, data=filterData, ylim=c(3,8),xlim=c(4,8), smooth=F, xbin=128)

##Plot FL2(mCherry) versus FSC 
xyplot(FL2.A ~ FSC.A, data=filterData, ylim=c(0,7),xlim=c(4,8), smooth=F, xbin=128)

##Plot FL1 (GFP) versus FL2 (Red)
xyplot(FL1.A ~ FL2.A, data=filterData, ylim=c(0,8),xlim=c(0,6), smooth=F, xbin=128)

```


###Step 5: Quantify fluorescence for each sample and plot a histogram

diagnostic values can be defined for plotting purposes
###Step 5: Quantify fluorescence for each sample and plot a histogram diagnostic values can be defined for plotting purposes
```{r}
#define critical values that can superimposed on plots for easy visual comparison

gfp.bg <- 3.9  #a background value for GFP
gfp.wt <- 5.9 #a value for wildtype GFP expression
red.bg <- 3.03 #a background value for the red channel
red.wt <- 3.75 #a value for wildtype Red expression
haploid.fsc <- 6.43 #an empirical value for forward scatter for haploids
diploid.fsc <- 6.54 #an empirical value for forward scatter for diploids
gfp.norm <- 0.935 #an empricial value for gfp expression normalized by forward scatter
red.norm <- 0.57 #an empricial value for red expression normalized by forward scatter
gfp.red.norm <- 1.5 #an empricial value for gfp expression normalized by red channel

```


```{r}
#record summary statistics in a matrix 
summary.stats <- matrix(data = NA, nrow = length(filterData), ncol = 18, dimnames = list(sampleNames(filterData),c("FSC_median","FSC_mean", "FSC_sd","FL1_median", "FL1_mean","FL1_sd","normalizedGFP_median", "normalizedGFP_mean", "normalizedGFP_sd","FL2_median","FL2_mean","FL2_sd","normalizedRed_median","normalizedRed_mean", "normalizedRed_sd","GFPnormalizedByRed_median", "GFPnormalizedByRed_mean","GFPnormalizedByRed_sd")))

#retain ratio of GFP(i.e. FL1.A) to FSC.A for generating boxplot as well as FSC and FL1A 
total <- min(fsApply(filterData, each_col, length)) #use the sample containing the minimum number of points after size-based filtering to define the size of the matrix 

print(total)

comparison.FSC <- matrix(data = NA, nrow = total, ncol = length(filterData), byrow = FALSE,dimnames = NULL)
comparison.FL1 <- matrix(data = NA, nrow = total, ncol = length(filterData), byrow = FALSE,dimnames = NULL)
comparison.FL2 <- matrix(data = NA, nrow = total, ncol = length(filterData), byrow = FALSE,dimnames = NULL)
comparison.FL1NormFsc <- matrix(data = NA, nrow = total, ncol = length(filterData), byrow = FALSE,dimnames = NULL)
comparison.FL2NormFsc <- matrix(data = NA, nrow = total, ncol = length(filterData), byrow = FALSE,dimnames = NULL)
comparison.FL1NormFL2 <- matrix(data = NA, nrow = total, ncol = length(filterData), byrow = FALSE,dimnames = NULL)
#analyzed.samples <- data.frame(matrix(data = NA, nrow = dim(fsApply(flowData, each_col, length))[1], ncol = 5, byrow = TRUE,dimnames = NULL))
analyzed.samples <- data.frame("Well" = character(0), "Strain" = character(0), "Genotype"=character(0), "Ploidy"=character(0), "Media"=character(0) )
#for each sample plot a histogram of the normalized data, raw FSC and raw GFP per row
par(mfrow=c(1,2), mar=c(5.1,2.1,2.1,2.1), oma=c(1.5,2,1,1))

#extract data from flowFrames to plot histograms of values and record summary statistics
for (i in 1:length(filterData)){
 
  temp <- exprs(filterData[[i]]) #exprs() extracts a matrix of the values from the flowframe
 
  sample.info <- sample.sheet[sample.sheet$Well==unlist(strsplit(sampleNames(filterData)[i], "[.]"))[1],] #get the info from the sample sheet corresponding to the well name.
  analyzed.samples <-  rbind(analyzed.samples,sample.info)
  
  ##########################################
  #record summary statistics for the sample#
  ##########################################
  
  #FSC
  summary.stats[i,1] <- median(temp[,1]) 
  summary.stats[i,2] <-mean(temp[,1])  
  summary.stats[i,3] <- sd(temp[,1])
  #FL1
  summary.stats[i,4] <- median(temp[,3])
  summary.stats[i,5] <-mean(temp[,3])  
  summary.stats[i,6] <- sd(temp[,3])
  #FL1 (GFP) divided by FSC
  summary.stats[i,7] <- median(temp[,3]/temp[,1])
  summary.stats[i,8] <-mean(temp[,3]/temp[,1])  
  summary.stats[i,9] <- sd(temp[,3]/temp[,1])
  #FL2
  summary.stats[i,10] <- median(temp[,4])
  summary.stats[i,11] <-mean(temp[,4])  
  summary.stats[i,12] <- sd(temp[,4])
  #FL2 (Red) divided by FSC
  summary.stats[i,13] <- median(temp[,4]/temp[,1])
  summary.stats[i,14] <-mean(temp[,4]/temp[,1])  
  summary.stats[i,15] <- sd(temp[,4]/temp[,1])
  #FL1 (GFP) divided by FL2 (Red)
  summary.stats[i,16] <- median(temp[,3]/temp[,4])
  summary.stats[i,17] <-mean(temp[,3]/temp[,4])  
  summary.stats[i,18] <- sd(temp[,3]/temp[,4])  
  
  ##############################################
  #plot histograms of the channels of interest##
  ##############################################

  ###############
  #Green channel#
  ###############
  
  #FL1 (GFP)
  hist(temp[,3], br=500, xlab = "FL1", main = "FL1") 
  abline(v=gfp.bg, col="yellow", lty=2, lwd=2)
  abline(v=gfp.wt, col="green", lty=2, lwd=2) 
  legend("topleft",  legend=paste("median FL1 = ",round(median(temp[,3]), digits=4),sep=""))

  #GFP divided by FSC
  hist(temp[,3]/temp[,1], br=500, xlim=c(0,1.5), xlab = "FL1/FSC", main = "FL1/FSC") 
  abline(v=gfp.norm, col="green", lty=2, lwd=2 )
  legend("topleft",  legend=paste("median GFP / FSC=",round(median(temp[,3]/temp[,1]), digits=4),sep=""))
  
  mtext(paste("Well = ", sample.info[1,1], " ; Strain = ", sample.info[1,2], " ; Genotype = ", sample.info[1,3], " ; Ploidy = ", sample.info[1,4], " ; Media = ", sample.info[1,5], sep=""), outer = TRUE, cex = 1.0)
  
  ###############
  #Red channel#
  ###############
  #FL2 (Red)
  hist(temp[,4], br=500, xlab = "FL2", main = "FL2", xlim=c(1,8)) 
  abline(v=red.bg, col="yellow", lty=2, lwd=2)
  abline(v=red.wt, col="red", lty=2, lwd=2) 
  legend("topleft",  legend=paste("median FL2=",round(median(temp[,4]), digits=4),sep=""))
 
  #FL2 divided by FSC
  hist(temp[,4]/temp[,1], br=500, xlim=c(0,1.5), xlab = "FL2/FSC", main = "FL2/FSC") 
  abline(v=red.norm, col="red", lty=2, lwd=2 )
  legend("topleft",  legend=paste("median FL2 / FSC=",round(median(temp[,4]/temp[,1]), digits=4),sep=""))

  mtext(paste("Well = ", sample.info[1,1], " ; Strain = ", sample.info[1,2], " ; Genotype = ", sample.info[1,3], " ; Ploidy = ", sample.info[1,4], " ; Media = ", sample.info[1,5], sep=""), outer = TRUE, cex = 1.0)
  
  ###############
  #Other#########
  ###############
  
  #FL1 divided by FL2
  hist(temp[,4]/temp[,3], br=500, xlim=c(0,1.5), xlab = "FL2/FL1", main = "FL1/FL2") 
  abline(v=gfp.red.norm, col="purple", lty=2, lwd=2)
  legend("topleft",  legend=paste("median FL1 / FL2=",round(median(temp[,4]/temp[,3]), digits=4),sep=""))

    #FSC
  hist(temp[,1], br=500, xlab = "FSC", main = "FSC", xlim=c(4,8)) 
  abline(v=haploid.fsc, col="blue", lty=2, lwd=2)
  abline(v=diploid.fsc, col="grey", lty=2, lwd=2)
  legend("topleft",  legend=paste("median FSC=",round(median(temp[,1]), digits=4),sep=""))
  
  mtext(paste("Well = ", sample.info[1,1], " ; Strain = ", sample.info[1,2], " ; Genotype = ", sample.info[1,3], " ; Ploidy = ", sample.info[1,4], " ; Media = ", sample.info[1,5], sep=""), outer = TRUE, cex = 1.0)

print("-------------------------------------------------------")
print("-----------------------------------")
print("----------------------")

  ############################################################
  #keep the data set for generating boxplots comparing values#
  ############################################################
  
  #Note that the amount of data kept for each sample is defined by the lowest count among all the samples.
  comparison.FSC[1:total,i] <- temp[1:total,1] #FSC
  comparison.FL1[1:total,i] <- temp[1:total,3] #FL1 (GFP)
  comparison.FL1NormFsc[1:total,i] <- temp[1:total,3]/temp[1:total,1] #GFP/FSC
  comparison.FL2[1:total,i] <- temp[1:total,4] #FL2 
  comparison.FL2NormFsc[1:total,i] <- temp[1:total,4]/temp[1:total,1] #FL2/FSC
  comparison.FL1NormFL2[1:total,i] <- temp[1:total,3]/temp[1:total,4] #FL1/FL2
  
}

par(mfrow=c(1,1)) #change number of plots per row back to standard
```


###Step 6: Generate boxplot comparing expression values between samples and print table

```{r}
par(mar=c(8.1,4.1,4.1,2.1)) #create more space at lower margin
#boxplot(comparison.FSC, names=sampleNames(filterData), notch = TRUE, col = "gray", ylab="FSC", cex.axis=0.5,las=2, outline=F)
j <- 2 #desired label: 1 = Well, 2 = strain, 3 = genotype, 4 = ploidy, 5 = media

boxplot(comparison.FSC, names=analyzed.samples[,j], notch = TRUE, col = "gray", ylab="FSC", cex.axis=0.5,las=2, outline=F)
abline(h=haploid.fsc, lty=2, col=2)
abline(h=diploid.fsc, lty=2, col=3)

boxplot(comparison.FL1, names=analyzed.samples[,j], notch = TRUE, col = "lightgreen", ylab="FL1", cex.axis=0.5,las=2, outline=F)
abline(h=gfp.bg ,lty=2, lwd=3, col="yellow")
abline(h=gfp.wt, lty = 2, lwd=3, col="green")

boxplot(comparison.FL1NormFsc, names=analyzed.samples[,j], notch = TRUE, col = "green", ylab="FL1/FSC", cex.axis=0.5,las=2, outline=F)
abline(h=gfp.norm, lty=2, lwd=3, col="blue")

boxplot(comparison.FL2, names=analyzed.samples[,j], notch = TRUE, col = "pink", ylab="FL2", cex.axis=0.5,las=2, outline=F)
abline(h=red.bg, lty=2, lwd=3, col="pink")
abline(h=red.wt, lty=2, lwd=3, col="red")

boxplot(comparison.FL2NormFsc, names=analyzed.samples[,j], notch = TRUE, col = "red", ylab="FL2/FSC", cex.axis=0.5,las=2, outline=F)
abline(h=red.norm, lty=2, lwd=3, col="red")

boxplot(comparison.FL1NormFL2, names=analyzed.samples[,j], notch = TRUE, col = "purple", ylab="FL1/FL2", cex.axis=0.5,las=2, outline=F)
abline(h=gfp.red.norm, lty=2, lwd=3, col="purple")

par(mar=c(5.1,4.1,4.1,2.1)) #reset margins to default

#generate a summary table containing all the recorded statistics

summary.stats <- cbind(analyzed.samples, summary.stats)
print(summary.stats)


```

###Step 7: Save plots or tables for use outside of R

```{r}
#export complete data table
write.table(summary.stats, file="Data_Summary.txt", row.names=TRUE, quote=F, sep="\t")

#specify which column of datasheet to include as sample name
j <- 2

#print boxplot of FSC
pdf(file="Boxplot_FSC.pdf", height=8, width=12)
par(mar=c(8.1,4.1,4.1,2.1)) #create more space at lower margin
boxplot(comparison.FSC, names=analyzed.samples[,j], notch = TRUE, col = "gray", ylab="FSC", cex.axis=0.6,las=2, outline=F)
abline(h=haploid.fsc, lty=2, col=2)
abline(h=diploid.fsc, lty=2, col=3)
dev.off()
par(mar=c(5.1,4.1,4.1,2.1)) #reset margins to default

#print boxplot of FL1
pdf(file="Boxplot_FL1.pdf", height=8, width=12)
par(mar=c(8.1,4.1,4.1,2.1)) #create more space at lower margin
boxplot(comparison.FL1, names=analyzed.samples[,j], notch = TRUE, col = "blue", ylab="FL1", cex.axis=0.6,las=2, outline=F)
abline(h=gfp.bg ,lty=2, lwd=3, col="yellow")
abline(h=gfp.wt, lty = 2, lwd=3, col="green")
dev.off()
par(mar=c(5.1,4.1,4.1,2.1)) #reset margins to default

#print boxplot of FL1 normalized by forward scatter
pdf(file="Boxplot_FSCNormalizedGFP.pdf", height=8, width=12)
par(mar=c(8.1,4.1,4.1,2.1)) #create more space at lower margin
boxplot(comparison.FL1NormFsc, names=analyzed.samples[,j], notch = TRUE, col = "green", ylab="GFP normalized by FSC", cex.axis=0.6,las=2, outline=F)
abline(h=gfp.norm, lty=2, lwd=3, col="blue")
dev.off()
par(mar=c(5.1,4.1,4.1,2.1)) #reset margins to default

#print boxplot of FL2
pdf(file="Boxplot_FL2.pdf", height=8, width=12)
par(mar=c(8.1,4.1,4.1,2.1)) #create more space at lower margin
boxplot(comparison.FL2, names=analyzed.samples[,j], notch = TRUE, col = "pink", ylab="FL2", cex.axis=0.6,las=2, outline=F)
abline(h=red.bg, lty=2, lwd=3, col="pink")
abline(h=red.wt, lty=2, lwd=3, col="red")
dev.off()
par(mar=c(5.1,4.1,4.1,2.1)) #reset margins to default

#print boxplot of normalized FL2
pdf(file="Boxplot_FSCnormalizedFL2.pdf", height=8, width=12)
par(mar=c(8.1,4.1,4.1,2.1)) #create more space at lower margin
boxplot(comparison.FL2NormFsc, names=analyzed.samples[,j], notch = TRUE, col = "red", ylab="FL2/FSC", cex.axis=0.6,las=2, outline=F)
abline(h=red.norm, lty=2, lwd=3, col="red")
dev.off()
par(mar=c(5.1,4.1,4.1,2.1)) #reset margins to default

#print boxplot of FL1 normalized by FL2
pdf(file="Boxplot_FL1normalizedFL2.pdf", height=8, width=12)
par(mar=c(8.1,4.1,4.1,2.1)) #create more space at lower margin
boxplot(comparison.FL1NormFL2, names=analyzed.samples[,j], notch = TRUE, col = "purple", ylab="FL1/FL2", cex.axis=0.6,las=2, outline=F)
abline(h=gfp.red.norm, lty=2, lwd=3, col="purple")
dev.off()
par(mar=c(5.1,4.1,4.1,2.1)) #reset margins to default

# # #example of how to to generate a pdf of a histogram
# j <- 2 #change index depending on which sample you want
# 
# temp <- exprs(filterData[[j]]) 
# pdf(file="Example_histogram.pdf", height=6, width=6)
# hist(temp[,3]/temp[,1], br=500, xlim=c(0,1.5), xlab = "Normalized expression signal (GFP/FSC)", main =  sampleNames(filterData)[j]) 
#   legend("topleft",  legend=paste("median=",round(median(temp[,3]/temp[,1]), digits=4),sep=""))
# dev.off()

```
